---
title: "AHS Overlap Population Fibromyalgia Study"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "janitor", "broom", "haven", "tableone", "gtsummary")
sapply(pacs, require, character.only = TRUE)

# Read SAS data
# n.obs = 5067
overlap0 <- read_sas("./data/choi_master.sas7bdat") %>% 
  clean_names()

# Overlap population, female only -----------------------------------------

# n = 3136
overlap_f <- overlap0 %>% 
  filter(sex == 0) %>% 
  
  # Define incident cases of fibromyalgia
  mutate(
    fibro_inc = case_when(
      (fibroy >= 1 & fibroy <= 4) | fibro == 2 ~ 1,
      TRUE ~ 0
      ),
    fm_stat = factor(fibro_inc, labels = c("Non-case", "Case"))
  ) %>% 

  # Demographics and lifestyles
  mutate(
    agecat = cut(agein, breaks = c(20, 30, 40, 50, 60, Inf), right = FALSE),
    agecat = factor(agecat, labels = c("20-29", "30-39", "40-49", "50-59", "60+")),
    
    bmicat = cut(ahs1_bmi, breaks = c(0, 25, 30, Inf), right = FALSE),
    bmicat = factor(bmicat, labels = c("<25", "25-<30", "30+")),
      
    educat3 = recode(educcq, 1, 1, 1, 2, 3, 3),
    educat3 = factor(educat3, labels = c("Less than HS", "Some college", "College grad+")),
    
    marital3 = recode(maritlcq, 1, 2, 3, 3),
    marital3 = factor(marital3, labels = c("Not married", "Married", "Wid/Div/Sep")),
    
    smoke2 = recode(smoke, 1, 1, 0),
    smoke2 = factor(smoke2, labels = c("Never", "Ever")),
    
    employ2 = factor(workpay, labels = c("No", "Yes"))
  ) %>% 
  
  # Parenting
  mutate(
    parent_cold = case_when(
      (mcold == 1 & fcold == 0) | (mcold == 0 & fcold == 1) ~ 1,
      mcold == 1 & fcold == 1 ~ 2,
      TRUE ~ 0
    ),
    parent_cold = factor(parent_cold, labels = c("None", "One", "Both")),
    
    parent_warm = case_when(
      (mwarm == 1 & fwarm == 0) | (mwarm == 0 & fwarm == 1) ~ 1,
      mwarm == 1 & fwarm == 1 ~ 2,
      TRUE ~ 0
    ),
    parent_warm = factor(parent_warm, labels = c("None", "One", "Both")),
    
    # Parent situation -- Need to check
    # parent_situ = case_when(
    #   raised %in% 1:2 ~ 0,
    #   e2q9r1 == 1 | e2q9r2 == 1 | e2q9r3 == 1 | e2q9r4 == 1 ~ 1
    # ),
    # parent_situ = factor(parent_situ, labels = c("Married in home", "One died/sep/div")),
    
    # Raised by -- Need to check
    fam_struct = case_when(
      raised == 1 ~ 0,
      raised == 2 ~ 1,
      raised %in% 3:4 ~ 2,
      raised == 5     ~ 3
    ),
    fam_struct = factor(fam_struct, labels = c("Two birthparents", "Two parents*", "Single-birthparent", "Other")),
    
    # Personality traits
    # Note reversed signs -- Need to check
    urgency     = (11 - rushed) + competv + tasks + (11 - fasteat),
    urgency4    = cut(urgency, breaks = c(4, 15, 26, 30, Inf), right = FALSE),
    urgency4    = factor(urgency4, labels = c("4-14", "15-25", "26-29", '30+')),
    
    depression  = (2 - overcome) + (2 - happy) + (family - 1),
    depression4 = factor(depression),
    
    authority   = (2 - respect) + (2 - counton) + (2 - showfel),
    authority4  = factor(authority),
    
    hostility   = case_when(
      geteven == 2 & dislike == 1 ~ 0,
      (geteven == 1 & dislike == 1) | (geteven == 2 & dislike == 2) ~ 1, 
      geteven == 1 & dislike == 2 ~ 2,
    ),
    hostility3  = factor(hostility),
    
    jobstress = case_when(
      jobsat %in% 1:2 | jobfrus %in% 3:4 ~ 0,
      jobsat %in% 3:4 & jobfrus %in% 1:2 ~ 1
    ),
    # jobstress = ifelse(is.na(jobsat) | is.na(jobfrus), NA_integer_, jobstress),
    jobstress = factor(jobstress, labels = c("Low frus or hi satis", "Hi frus & low satis"))
  )

```

## Datasets 

* The overlap population between AHS-1 and AHS-2
  * See the manuscript how record linkage was done

* Contains 5067 subjects. Among them, there were n = 3136 women.

## Outcome

* Self-reported diagnosis of fibromyalgia from **AHS-2 baseline** questionnaire ([Page A4, under the section of musculoskeletal system](https://wiki.ahs2.org/_media/baseline:ahs_baseline_v3_-_page_a4.jpg))
  * Years since first diagnosis: Less than 5 years ago, 5-9 yrs ago, 10-14 yrs ago, 20+ yrs ago
  * Have you been treated for this in the last 12 months?: No or Yes
  * See the crosstab below
    * 127 females reported their FM diagnosis. None of these were diagnosed 20+ years ago
    * 64 females were treated in the last 12 month. Among them, 9 females did not indicate their diagnosis year
  * [**Need to check**] If diagnosed < 20 years ago or treated within the last 12 months, they were considered as incident cases
    * According to this classification rule, there are 136 incident cases (4.3%) out of 3136 women

```{r fibro_crosstab, echo = FALSE}

# Crosstab between fibroy and fibro
overlap_f %>%
  mutate(
    fibroy = factor(fibroy, labels = c("<5 yrs", "5-9 yrs", "10-14 yrs", "15-19 yrs", "20+ yrs"), levels = 1:5),
    fibro  = factor(fibro,  labels = c("No", "Yes"))
    ) %>% 
  tabyl(fibroy, fibro) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  knitr::kable()

```

## Exposure of interest

* All of the followings were derived from **AHS-1 questionnaire**, except for family structure in early life. Variable names in the data were indicated below
  * See also: Table 1 of the manuscript

* Early life experiences (See [the questions in AHS-1](./images/AHS1_Q37.png))
  * Warm parenting: `mwarm`, `fwarm`
  * Cold parenting: `mcold`, `fcold`
  * ~~Parent situation~~: Removed or replaced with "family structure" 
  * [**Need to check**] Family structure
    * Replace "Raised by" variable
    * Based on AHS-2 BQ, Page E2, Q8, asking "Up through age 16 years, were you mostly raised with" (See [the questions in AHS-2](./images/AHS2_BQ_E2Q8.png))
    * If answered "Your two birth parents", then categorized into "Two birthparents"
    * If answered "Two parents, but one or both...", then categorized into "Two parents*" (one or both were not birth parent)
    * If answered "A female birthparent only" or "A male birthparent only", then categorized into "Single-birthparent"
    * If answered "Other" then categorized into "Other"

* Psychologic characteristics (See [the questions in AHS-1](./images/AHS1_Q52.png))
  * Depression index score:
    * `family`, `happy`, `overcome`
    * yes/no
    * A negative response gets 1, otherwise 0
    * Points were summed up, ranging from 0 to 3
    
  * Hostility index score:
    * `dislike`, `geteven`
    * yes/no or agree/disagree
    * A negative response gets 1, otherwise 0
    * Points were summed up, ranging from 0 to 2
    
  * Authoritarian index score:
    * `counton`, `showfel`, `respect`
    * agree/disagree
    * A negative response gets 1, otherwise 0
    * Points were summed up, ranging from 0 to 3
    
  * Time urgency index score:  (See [the questions in AHS-1](./images/AHS1_Q61.png))
    * `rushed`, `competv`, `tasks`, `fasteat`
    * 10-point likert scale
    * Scales were reversed for `rushed` and `fasteat` (see AHS-1 questionnaire) before summing up
    * Possible range: from 4 to 40
    * [**Need to check**] The distribution of this score does **NOT** match with those shown in Table 3 of the manuscript

* Job stress
  * Based on AHS-1 Q21 and Q22 (See [the questions in AHS-1](./images/AHS1_Q21.png))
    * `jobsat`: How well satisfied are you with job
    * `jobfrus`: How often are you irritated or frustrated by your job
    * If answered ("Not too satisfied" OR "Not at all satisfied") AND ("Always" OR "Often"), then categorized to "High frustration and low satisfaction"
      * Otherwise categorized to "Low frustration or high satisfaction"

## Descriptive table

* All of the followings were derived from **AHS-1 questionnaire**, except for family structure (see the section above)

* See below for some of demographic/lifestyle variables:
  * Education: Based on the AHS-1 variable `EducCQ`, categorized into 3 levels as shown in the table
  * Employment: Based on AHS-1 Q20, categorized into 2 levels: Employed or Unemployed
    * Employed: Self-employed, full time or part time employed
    * Unemployed: Out of work, student, homemaker, volunteer worker, retired
  * Marital status: Based on the AHS-1 variable `MaritalCQ`, categorized into 3 levels as shown in the table
  * Smoking status: Based on AHS-1 Q27, categorized into 2 levels: Never or Ever

* [**Need to check**] Note the number of missing values as well. **How should we handle the missing values?** Any inclusion/exclusion criteria?

```{r descriptive_table, echo = FALSE, out.width = "50%"}

# Table 1 -----------------------------------------------------------------

# Variables to be included 
# It appears that all vars come from AHS-1
table_vars <- c(
  "agecat",
  "agein",
  "bmicat",
  "ahs1_bmi",
  "educat3",
  "employ2",
  "marital3",
  "smoke2",
  "parent_warm",
  "parent_cold",
  # "parent_situ",
  # "raised_by",
  "fam_struct",
  "depression4",
  "hostility3",
  "authority4",
  "urgency4",
  "jobstress"
)

overlap_f %>%
  select(all_of(table_vars), fm_stat) %>% 
  tbl_summary(
    by = "fm_stat",
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    digits = all_continuous() ~ 1,
    type = list(employ2 ~ "categorical"),
    missing = "always",
    missing_text = "(Missing)"
  ) %>%
  add_p(
    test = list(all_continuous() ~ "t.test"),
    pvalue_fun = label_style_pvalue(digits = 3),
  ) %>% 
  as_flex_table()

```

## Distribution for age and BMI

* See below for distributions of age and BMI:


```{r check_distribution, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 3}

overlap_f %>% 
  select(agein, ahs1_bmi) %>% 
  pivot_longer(1:2, names_to = "Variable", values_to = "Value") %>% 
  ggplot(aes(x = Value)) +
  geom_histogram(aes(y = after_stat(density)), fill = "gray") +
  geom_density(color = "cornflowerblue", linewidth = 1) +
  facet_wrap(~ Variable, scales = "free")

```

## Logistic models

* For each exposure variable of interest, first we fit logistic models as incident fibromyalgia as the outcome to obtain unadjusted odds ratios associated with the exposure variable


### Unadjusted odds ratios

* Unadjusted odds ratios for each exposure variable are shown below
  * Note that reference groups are not shown in the table
    * For warm parenting, the reference is "Both"
    * For other exposure variables, the reference is the first level shown in the descriptive table 
    
* There was a significant trend for warm/cold parenting, depression, hostility, authority

```{r unadjusted_odds_ratios, echo = FALSE, warning = FALSE, message = FALSE}

# Exposure of interest ----------------------------------------------------

# Exposure variables of interest
ind_vars <- c(
  "parent_warm",
  "parent_cold",
  # "parent_situ",
  # "raised_by",
  "fam_struct",
  "depression4",
  "hostility3",
  "authority4",
  "urgency4",
  "jobstress"
)

# Covariates
covars <- c(
  "agein",
  "ahs1_bmi",
  "educat3",
  "employ2",
  "marital3"
)

# Logistic regression: Unadjusted ORs -------------------------------------

# Reverse order for parent_warm
overlap_f2 <- overlap_f %>% 
  mutate(parent_warm = fct_rev(parent_warm))

# Unadjusted ORs
# Create models
model_formulas <- map(ind_vars, ~reformulate(.x, response = "fibro_inc")) %>% 
  setNames(ind_vars)

# Run models
unadjusted_results <- map(model_formulas, ~glm(.x, data = overlap_f2, family = "binomial"))

# Get unadjusted ORs
unadj_OR <- map_dfr(unadjusted_results, tidy, .id = "predictor") %>%
  mutate(
    odds_ratio = exp(estimate), 
    conf_low = exp(estimate - 1.96 * std.error),
    conf_high = exp(estimate + 1.96 * std.error)
  ) %>%
  filter(term != "(Intercept)")

unadj_OR <- unadj_OR %>%  
  mutate(predictor = factor(predictor, levels = unique(unadj_OR$predictor))) %>% 
  select(predictor, term, odds_ratio, conf_low, conf_high, p.value) 

# Trend p-values
ind_vars_num <- paste0("as.numeric(", ind_vars, ")")
model_formulas <- map(ind_vars_num, ~reformulate(.x, response = "fibro_inc")) %>% 
  setNames(ind_vars)

trend_p <- model_formulas %>% 
  map(~ glm(.x, data = overlap_f2, family = "binomial")) %>% 
  map_dbl(~ coef(summary(.x))[2, "Pr(>|z|)"])

trend_p_df <- data.frame(trend.p = trend_p) %>% 
  rownames_to_column("predictor") %>% 
  filter(predictor != "jobstress") %>% 
  mutate(trend.p = format.pval(trend.p, digits = 2))

options(knitr.kable.NA = "")

unadj_OR %>% 
  left_join(trend_p_df, by = "predictor") %>% 
  group_by(predictor) %>%
  mutate(across(odds_ratio:conf_high, round, 2)) %>%
  mutate(
    p.value = round(p.value, 4),
    trend.p = if_else(row_number() == 1, trend.p, NA_character_)
    ) %>%
  ungroup() %>% 
  knitr::kable()

```

### Adjusted odds ratios

* This time, we fit multivariable logistic models adjusting for:
  * Age as continuous
  * BMI as continuous
  * Education as categorical
  * Employment as binary (no/yes)
  * Marital status as categorical

* Adjusted odds ratios for each exposure variable are shown below
  * Note that reference groups are not shown in the table
    * For warm parenting, the reference is "Both"
    * For other exposure variables, the reference is the first level shown in the descriptive table 
    
* There was a significant trend for warm/cold parenting, depression, hostility, authority

```{r adjusted_odds_ratios, echo = FALSE, warning = FALSE, message = FALSE}

# Logistic regression: Adjusted ORs ---------------------------------------

# Adjusted ORs
# Create models
model_formulas <- map(ind_vars, ~reformulate(c(.x, covars), response = "fibro_inc")) %>% 
  setNames(ind_vars)

# Run models
adjusted_results <- map(model_formulas, ~glm(.x, data = overlap_f2, family = "binomial"))

# Get adjusted ORs
adjusted_OR <- map_dfr(adjusted_results, tidy, .id = "predictor") %>%
  mutate(
    odds_ratio = exp(estimate), 
    conf_low = exp(estimate - 1.96 * std.error),
    conf_high = exp(estimate + 1.96 * std.error)
  ) %>%
  filter(term != "(Intercept)") %>% 
  select(predictor, term, odds_ratio, conf_low, conf_high, p.value)  
  
# Trend p-values
ind_vars_num <- paste0("as.numeric(", ind_vars, ")")
model_formulas <- map(ind_vars_num, ~reformulate(c(.x, covars), response = "fibro_inc")) %>% 
  setNames(ind_vars)

trend_p <- model_formulas %>% 
  map(~ glm(.x, data = overlap_f2, family = "binomial")) %>% 
  map_dbl(~ coef(summary(.x))[2, "Pr(>|z|)"])

trend_p_df <- data.frame(trend.p = trend_p) %>% 
  rownames_to_column("predictor") %>% 
  filter(predictor != "jobstress") %>% 
  mutate(trend.p = format.pval(trend.p, digits = 2))

adjusted_OR %>% 
  filter(grepl(paste0(ind_vars, collapse = "|"), term)) %>% 
  left_join(trend_p_df, by = "predictor") %>% 
  group_by(predictor) %>%
  mutate(across(odds_ratio:conf_high, round, 2)) %>%
  mutate(
    p.value = round(p.value, 4),
    trend.p = if_else(row_number() == 1, trend.p, NA_character_)
    ) %>%
  ungroup() %>% 
  knitr::kable()

```

