---
title: "AHS Overlap Population Fibromyalgia Study"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Required packages
pacs <- c("tidyverse", "janitor", "broom", "haven", "tableone")
sapply(pacs, require, character.only = TRUE)

# Read SAS data
# n.obs = 5067
overlap0 <- read_sas("./data/choi_master.sas7bdat") %>% 
  clean_names()

# Overlap population, female only -----------------------------------------

# n = 3136
overlap_f <- overlap0 %>% 
  filter(sex == 0) %>% 
  
  # Define incident cases of fibromyalgia
  mutate(
    fibro_inc = case_when(
      (fibroy >= 1 & fibroy <= 4) | fibro == 2 ~ 1,
      TRUE ~ 0
      ),
    fm_stat = factor(fibro_inc, labels = c("Non-case", "Case"))
  ) %>% 

  # Demographics and lifestyles
  mutate(
    agecat = cut(agein, breaks = c(20, 30, 40, 50, 60, Inf), right = FALSE),
    agecat = factor(agecat, labels = c("20-29", "30-39", "40-49", "50-59", "60+")),
    
    bmicat = cut(ahs1_bmi, breaks = c(0, 25, 30, Inf), right = FALSE),
    bmicat = factor(bmicat, labels = c("<25", "25-<30", "30+")),
      
    educat3 = recode(educcq, 1, 1, 1, 2, 3, 3),
    educat3 = factor(educat3, labels = c("Less than HS", "Some college", "College grad+")),
    
    marital3 = recode(maritlcq, 1, 2, 3, 3),
    marital3 = factor(marital3, labels = c("Not married", "Married", "Wid/Div/Sep")),
    
    smoke2 = recode(smoke, 1, 1, 0),
    smoke2 = factor(smoke2, labels = c("Never", "Ever")),
    
    employ2 = factor(workpay, labels = c("No", "Yes"))
  ) %>% 
  
  # Parenting
  mutate(
    parent_cold = case_when(
      (mcold == 1 & fcold == 0) | (mcold == 0 & fcold == 1) ~ 1,
      mcold == 1 & fcold == 1 ~ 2,
      TRUE ~ 0
    ),
    parent_cold = factor(parent_cold, labels = c("None", "One", "Both")),
    
    parent_warm = case_when(
      (mwarm == 1 & fwarm == 0) | (mwarm == 0 & fwarm == 1) ~ 1,
      mwarm == 1 & fwarm == 1 ~ 2,
      TRUE ~ 0
    ),
    parent_warm = factor(parent_warm, labels = c("None", "One", "Both")),
    
    # Parent situation -- Need to check
    parent_situ = case_when(
      raised %in% 1:2 ~ 0,
      e2q9r1 == 1 | e2q9r2 == 1 | e2q9r3 == 1 | e2q9r4 == 1 ~ 1
    ),
    parent_situ = factor(parent_situ, labels = c("Married in home", "One died/sep/div")),
    
    # Raised by -- Need to check
    raised_by = case_when(
      raised %in% 1:2 ~ 0,
      raised %in% 3:4 ~ 1,
      raised == 5     ~ 2
    ),
    raised_by = factor(raised_by, labels = c("Both", "One", "Non-parents")),
    
    # Personality traits
    # Note reversed signs -- Need to check
    urgency     = (11 - rushed) + competv + tasks + (11 - fasteat),
    urgency4    = cut(urgency, breaks = c(4, 15, 26, 30, Inf), right = FALSE),
    urgency4    = factor(urgency4, labels = c("4-14", "15-25", "26-29", '30+')),
    
    depression  = (2 - overcome) + (2 - happy) + (family - 1),
    depression4 = factor(depression),
    
    authority   = (2 - respect) + (2 - counton) + (2 - showfel),
    authority4  = factor(authority),
    
    hostility   = case_when(
      geteven == 2 & dislike == 1 ~ 0,
      (geteven == 1 & dislike == 1) | (geteven == 2 & dislike == 2) ~ 1, 
      geteven == 1 & dislike == 2 ~ 2,
    ),
    hostility3  = factor(hostility),
    
    jobstress = case_when(
      jobsat %in% 1:2 | jobfrus %in% 3:4 ~ 0,
      jobsat %in% 3:4 & jobfrus %in% 1:2 ~ 1
    ),
    # jobstress = ifelse(is.na(jobsat) | is.na(jobfrus), NA_integer_, jobstress),
    jobstress = factor(jobstress, labels = c("Low frus or hi satis", "Hi frus & low satis"))
  )

```

## Datasets 

* The overlap population between AHS-1 and AHS-2
  * See the manuscript how record linkage was done

* Contains 5067 subjects. Among them, there were n = 3136 women.

## Outcome

* Self-reported diagnosis of fibromyalgia from **AHS-2 baseline** questionnaire ([Page A4, under the section of musculoskeletal system](https://wiki.ahs2.org/_media/baseline:ahs_baseline_v3_-_page_a4.jpg))
  * Years since first diagnosis: Less than 5 years ago, 5-9 yrs ago, 10-14 yrs ago, 20+ yrs ago
  * Have you been treated for this in the last 12 months?: No or Yes
  * See the crosstab below
    * 127 females reported their FM diagnosis. None of these were diagnosed 20+ years ago
    * 64 females were treated in the last 12 month. Among them, 9 females did not indicate their diagnosis year
  * [**Need to check**] If diagnosed < 20 years ago or treated within the last 12 months, they were considered as incident cases

```{r fibro_crosstab, echo = FALSE}

# Crosstab between fibroy and fibro
overlap_f %>%
  mutate(
    fibroy = factor(fibroy, labels = c("<5 yrs", "5-9 yrs", "10-14 yrs", "15-19 yrs", "20+ yrs"), levels = 1:5),
    fibro  = factor(fibro,  labels = c("No", "Yes"))
    ) %>% 
  tabyl(fibroy, fibro) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  knitr::kable()

```

## Exposure of interest

* All of the followings were derived from **AHS-1 questionnaire**, except for.... Variable names in the data were indicated below
  * See also: Table 1 of the manuscript

* Early life experiences (See [the questions in AHS-1](./images/AHS1_Q37.png))
  * Warm parenting: `mwarm`, `fwarm`
  * Cold parenting: `mcold`, `fcold`
  * Parent situation: 
  * Raised by:

* Psychologic characteristics (See [the questions in AHS-1](./images/AHS1_Q52.png))
  * Depression index score:
    * `family`, `happy`, `overcome`
    * yes/no
    * A negative response gets 1, otherwise 0
    * Points were summed up, ranging from 0 to 3
    
  * Hostility index score:
    * `dislike`, `geteven`
    * yes/no or agree/disagree
    * A negative response gets 1, otherwise 0
    * Points were summed up, ranging from 0 to 2
    
  * Authoritarian index score:
    * `counton`, `showfel`, `respect`
    * agree/disagree
    * A negative response gets 1, otherwise 0
    * Points were summed up, ranging from 0 to 3
    
  * Time urgency index score:  (See [the questions in AHS-1](./images/AHS1_Q61.png))
    * `rushed`, `competv`, `tasks`, `fasteat`
    * 10-point likert scale
    * Scales were reversed for `rushed` and `fasteat` (see AHS-1 questionnaire) before summing up
    * Possible range: from 4 to 40

## Descriptive table

```{r descriptive_table, echo = FALSE}

# Table 1 -----------------------------------------------------------------

# Variables to be included 
# It appears that all vars come from AHS-1
table_vars <- c(
  "agecat",
  "agein",
  "bmicat",
  "ahs1_bmi",
  "educat3",
  "employ2",
  "marital3",
  "smoke2",
  "parent_warm",
  "parent_cold",
  "parent_situ",
  "raised_by",
  "depression4",
  "hostility3",
  "authority4",
  "urgency4",
  "jobstress"
)

overlap_f %>% CreateTableOne(
  vars = table_vars, 
  strata = "fm_stat", 
  data = .
  ) %>% 
  print(showAllLevels = TRUE) %>% 
  kableone()

```

## Logistic models
